---
title: "Inflation"
format: 
  dashboard:
    #logo: 
    nav-buttons: [github]
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: attaching libraries
setwd("~/My Documents/inflation-dash")
library(tidyverse)
library(gt)
```

```{r}
#| label: loading data
data = read.csv("merged_data.csv")
data <- mutate(data, time = as.Date(time))

income_comps = c("contr_unit_labor_cost", "contr_unit_profit", "contr_unit_tax")
```

# Sidebar {.sidebar width="20%"}
<br>

This dashboard displays annual inflation based on GDP deflator and Consumer Price Index for a set of countries. It also displays the contribution of unit labor cost, unit profit, and unit tax to inflation measured through the percentage change in the GDP deflator for a given country.

Pick a country below

# Plots {orientation="rows"}

## Inflation: Deflator- and CPI-Based

```{r}
#| label: plot-decomposition
#| title: Decomposition of GDP Deflator


ref_area_examp = "FRA"

data |>
  filter(ref_area == ref_area_examp) |>
  filter(series %in% income_comps) |>
  ggplot(aes(x = time, y = value, fill = series)) +
  geom_bar(stat = "identity") +
  labs(x="", y="Percent") +
  geom_line(
    data = filter(data, ref_area == ref_area_examp, series == "inflation_def"), 
    aes(time, value, color = "inflation_def")) +
  scale_fill_viridis_d(
    breaks = c("contr_unit_labor_cost", "contr_unit_profit", "contr_unit_tax"),
    labels = c("Unit labor cost", "Unit profit", "Unit tax"),
    option = "turbo") +
  scale_color_manual(
    labels = "Deflator",
    values = c("inflation_def" = "red")
  ) +
  theme(
    legend.title = element_blank(),
    legend.position = "bottom"
  )
```

```{r}
#| title: Annual Inflation Based on GDP Deflator and CPI, Quarterly Data
data |>
  filter(series %in% c("inflation_def", "inflation_cpi")) |>
  filter(reference_area == "France") |>
  ggplot(aes(time, value, color = series)) +
  geom_line(size=0.7) +
  scale_color_brewer(
    breaks = c("inflation_def", "inflation_cpi"),
    labels = c("Deflator", "Consumer Price Index"),
    palette = "Dark2"
  ) +
  labs(y = "Percent", x = "") +
  theme(
    #legend.position = "inside",
    #legend.position.inside = c(0.5, 0.8),
    #legend.background = element_blank(),
    legend.title = element_blank(),
    legend.position = "bottom"
  )
```


## Decomposition



# Tables

```{r}
#| label: table-decomposition
#| title: ttt
data |>
  filter(reference_area == "France") |>
  filter(series %in% c("inflation_def", "inflation_cpi", income_comps)) |>
  pivot_wider(names_from = series, values_from = value) |>
  select(c(reference_area, time, contr_unit_labor_cost,
           contr_unit_profit, contr_unit_tax,
           inflation_def, inflation_cpi)) |>
  rename(
    Country = reference_area,
    Period = time,
    `Unit labor cost` = contr_unit_labor_cost,
    `Unit profit` = contr_unit_profit,
    `Unit tax` = contr_unit_tax,
    Deflator = inflation_def,
    CPI = inflation_cpi
  ) |>
  arrange(desc(Period)) |>
  gt() |>
  fmt_number(
    columns = 3:7,
    decimals = 2
  ) |>
  fmt_date(
    columns = Period,
    date_style = "year_quarter"
  )
```

Decadal averages goes here.