---
title: "Inflation"
format: 
  dashboard:
    #logo: 
    nav-buttons: [github]
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: attaching libraries
setwd("~/My Documents/inflation-dash")
library(tidyverse)
```

```{r}
#| label: loading data
data = read.csv("merged_data.csv")
data <- mutate(data, time = as.Date(time))

income_comps = c("contr_unit_labor_cost", "contr_unit_profit", "contr_unit_tax")
```

# Sidebar {.sidebar}
This dashboard displays annual inflation based on GDP deflator and Consumer Price Index for a set of countries. It also displays the contribution of unit labor cost, unit profit, and unit tax to inflation measured through the percentage change in the GDP deflator for a given country.

# Plots {orientation="rows"}

## Inflation: Deflator- and CPI-Based {height="40%"}

```{r}
#| title: Annual Inflation Based on GDP Deflator and CPI, Quarterly Data
data |>
  filter(series %in% c("inflation_def", "inflation_cpi")) |>
  filter(reference_area == "France") |>
  ggplot(aes(time, value, color = series)) +
  geom_line(size=0.7) +
  scale_color_viridis_d(
    breaks = c("inflation_def", "inflation_cpi"),
    labels = c("Deflator", "CPI")
  ) +
  labs(y = "Percent", x = "") +
  theme(
    #legend.position = "inside",
    #legend.position.inside = c(0.5, 0.8),
    #legend.background = element_blank(),
    legend.title = element_blank()
  )
```


```{r}
#| title: Decomposition of GDP Deflator
ref_area_examp = "FRA"
data |>
  filter(ref_area == ref_area_examp) |>
  filter(series %in% income_comps) |>
  ggplot(aes(x = time, y = value, fill = series)) +
  geom_col() +
  labs(x="", y="Percent") +
  geom_line(
    data = filter(data, ref_area == ref_area_examp, series == "inflation_def"), aes(time, value), color = "red"
  ) +
  scale_fill_viridis_d(
    breaks = c("contr_unit_labor_cost", "contr_unit_profit", "contr_unit_tax"),
    labels = c("Unit labor cost", "Unit profit", "Unit tax"),
    option = "turbo"
  ) +
  theme(
    legend.title = element_blank()
  )
```

## Data


# Tables